Generalized Bradley-Terry models

### 2025-03-31 inst/jss_hyper3_real.Rnw
typos

About 22\% of matches in Table~\ref{oldfirm} end in a draw and
\pkg{hyper2} formalism may be used to assess this.  Consider the
likelihood function described in Table~\ref{windrawloss}; this is not
easily represented as a generalized linear model using the techniques
of~\citet{turner2020}.  Relatedlikelihood functions for drawn matches
were considered using the reified entity technique
of~\citet{hankin2020_jss} in which a drawn match was considered to be
won by a fictitious entity (a ``draw monster'') who possessed his own
BT strength.  However, the likelihood function of
table~\ref{windrawloss} has a number of desirable featureas not
present in the reified entity approach of~\citet{hankin2020_jss}.
Firstly, if we consider two players of identical strengths and nullify
the home ground effect (that is, $\lambda=1$) then the probability of
a draw is $\frac{2D}{2+2D}$; the odds of a draw is simply $D$,
regardless of team strength.  Secondly, considering only won (that is,
not score-drawn) games, the probability of a home win is
$\frac{\lambda}{1+\lambda}$, that is, the odds of a home win is simply
$\lambda$, independently of the player strength. Such arguments give
direct operational significance to $\lambda$ and $D$.

@@

Equation~\ref{mostgeneral} is very general and may be applied to a
range of situations in which standard Bradley-Terry strengths are
modified in some way.  In the following section, I apply it to
Plackett-Luce \citep{plackett1975_jss} likelihood functions for
rank data

>>>>

About 22\% of matches in Table~\ref{oldfirm} end in a draw and
\pkg{hyper2} formalism may be used to assess this.  Consider the
likelihood function described in Table~\ref{windrawloss}; this is not
easily represented as a generalized linear model using the techniques
of~\citet{turner2020}.  Relatedlikelihood functions for drawn matches
were considered using the reified entity technique
of~\citet{hankin2020_jss} in which a drawn match was considered to be
won by a fictitious entity (a ``draw monster'') who possessed his own
BT strength.  However, the likelihood function of
table~\ref{windrawloss} has a number of desirable features not
present in the reified entity approach of~\citet{hankin2020_jss}.
Firstly, if we consider two players of identical strengths and nullify
the home ground effect (that is, $\lambda=1$) then the probability of
a draw is $\frac{2D}{2+2D}$; the odds of a draw is simply $D$,
regardless of team strength.  Secondly, considering only won (that is,
not score-drawn) games, the probability of a home win is
$\frac{\lambda}{1+\lambda}$, that is, the odds of a home win is simply
$\lambda$, independently of the player strength. Such arguments give
direct operational significance to $\lambda$ and $D$.

@@

Equation~\ref{mostgeneral} is very general and may be applied to a
range of situations in which standard Bradley-Terry strengths are
modified in some way.  In the following section, I apply it to
Plackett-Luce \citep{plackett1975_jss} likelihood functions for
rank data.
